---
title: "Process Steps Taken data"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading data


```{r read_data}
library(reshape2)
df <- read.csv("./activity.csv",header=TRUE)
```
## Calculate SUM of Steps Taken by day 
```{r data_adjust,echo=TRUE}
df1<-tapply(df$steps,factor(df$date), sum)
df1 <- as.data.frame(df1)
df1$date=rownames(df1)
names(df1) <- c("value","date")
dfMelt<- melt(df1,id=c("date"),measure.vars=c("value"))
```
## Calculate HISTO/Frequency of steps per day
```{r histo,echo=TRUE}
hist(dfMelt$value, xlab="Class of Total Number of Steps per day", ylab="Number of Days", main="Total Number of Steps taken each day")
```

## Calculate mean steps
```{r mean ,echo=TRUE}
mean<- mean(dfMelt$value,na.rm = TRUE)
print(mean)
```

## Calculate median steps
```{r median ,echo=TRUE}
median<- median(dfMelt$value,na.rm = TRUE)
print(median)
```
## Calculate pattern of average steps in day
```{r pattern ,echo=TRUE}
meanStepsbyInterval<-aggregate(df$steps,by=list(df$interval), FUN=mean,na.rm=TRUE)
names(meanStepsbyInterval) <- c("interval","meanSteps")
with(meanStepsbyInterval, plot(interval, meanSteps, type = "l"))
```

## Calculate interval with max steps
```{r max ,echo=TRUE}
maxInterval <-  meanStepsbyInterval[meanStepsbyInterval$meanSteps==max(meanStepsbyInterval$meanSteps),"interval"]
print(maxInterval)
```

## Calculate number of NAs
```{r NAs ,echo=TRUE}
na <- df[is.na(df$steps),"steps"]
print(length(na))
```

## Define df1 with NAs replaced by mean for that interval
```{r replace_NAs ,echo=FALSE}
fill_NA <- function(df) 
  {
   index <- 1
   df1 <- df
   meanSteps<-aggregate(df$steps,by=list(df$interval), FUN=mean,na.rm=TRUE)
   names(meanSteps) <- c("interval","meanSteps") 
   while(index <= length(df[,"steps"]))
     #while(index <= 20)
     {
    steps<- df[index,"steps"]
    #print(steps)
    interval<- df[index,"interval"]
   # print(interval)
    meanForInterval <- meanSteps[meanSteps$interval==interval,"meanSteps"]
    #print(meanForInterval)
    if (is.na(steps)) 
     {
     
       df1[index,"steps"] <- meanForInterval
       df1[index,"interval"] <- interval
       df1[index,"date"] <- df[index,"date"]
    }
    else  {
      df1[index,"steps"] <- steps
      df1[index,"interval"] <- interval
      df1[index,"date"] <- df[index,"date"]
    }
   index <- index+1
  
   }
   return(df1)
}

df1<- fill_NA(df)
str(df1)
```

## Histo of steps taken without NAs
```{r histo_noNas ,echo=TRUE}
sumStepsbyDate <- aggregate(df1$steps,by=list(df1$date),sum)
names(sumStepsbyDate) =c("date","sumSteps")
hist(sumStepsbyDate$sumSteps, xlab="Class of Total Number of Steps per day", ylab="Number of Days", main="Total Number of Steps taken each day w/o NAs")



```

## Calculate mean steps w/o NAs
```{r mean_NA ,echo=TRUE}
mean<- mean(sumStepsbyDate$sumSteps)
print(mean)
```

## Calculate median steps w/o NAs
```{r median_NA ,echo=TRUE}
median<- median(sumStepsbyDate$sumSteps)
print(median)
```


## Pattern of Steps in WorkDay and Holiday
```{r histo_workday ,echo=TRUE}
dayOfWeek <- function(df1) 
{
  index <- 1
  # df1 <- df
  
  while(index <= length(df1[,"steps"]))
    #while(index <= 20)
  {
    date<- df1[index,"date"]
    day <- weekdays(date)
    
    if (day=="domenica"|day=="sabato") 
    {
      df1[index,"weekday"] <- "holyDay"
     
    }
    else  {
     
      df1[index,"weekday"] <- "workDay"
    }
    index <- index+1
    
  }
  return(df1)
}
df1$date <- as.Date(df1$date,"%Y-%m-%d")
df1 <- dayOfWeek(df1 )
df1$weekday <- factor(df1$weekday)
meanSteps<-aggregate(df1$steps,by=list(df1$interval,df1$weekday), FUN=mean)
names(meanSteps) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, meanSteps, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```